---
title: "Discriminability"
author: "Eric Bridgeford"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{discriminability}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
    header-includes:
   - \usepackage{amsfonts}
   - \usepackage{amsmath}
   - \usepackage[linesnumbered,lined,boxed,commentsnumbered]{algorithm2e}
---

# Simple Example

In this example, we will have 2 populations of normally distributed data, where our variables are described:

\begin{align*}
  X_{1, i} \sim \mathcal{N}(\mu_1=0, \sigma_1 = .5) \\
  X_{2, i} \sim \mathcal{N}(\mu_2=0, \sigma_2 = .5)
\end{align*}

for $i=1:5$ dimensions, and samples $x_{v, j} \in \mathbb{R}^5$ for our $j^{th}$ realization of variable $v$:


```{r, fig.height=5, fig.width=6}
  require(fmriutils)
  require(MGC)
  n <- 2
  d <- 5
  s <- 2
  dat <- array(NaN, dim=c(n*s, d))
  mu1 <- 0
  mu2 <- 2
  sd1 <- .2
  sd2 <- .2
  dat[1,] <- rnorm(n=d, mean=mu1, sd=sd1)
  dat[2,] <- rnorm(n=d, mean=mu1, sd=sd1)
  dat[3,] <- rnorm(n=d, mean=mu2, sd=sd2)
  dat[4,] <- rnorm(n=d, mean=mu2, sd=sd2)
  fmriu.plot.plot_graph(t(dat), limits=c(-1, 3), xlabel = "dimension", ylabel="Index", legend.name="intensity", title="2 Simulated Observations of 2 Random Variables")
```

As we can see, our observations of the first variable in the first two vectors (1 and 2) are much closer to one another than to our observations of our second variable in the second pair of vectors (3 and 4) and vice versa for our two observations of our second variable in the second pair of vectors.

This can be easily confirmed with the distance matrix for our data:

```{r, fig.height=5, fig.width=6}
D <- discr.distance(dat)
fmriu.plot.plot_graph(D, title="Distance Matrix for Simulated Data", xlabel="Scan", ylabel="Scan", legend.name="distance", limits=c(min(D), max(D)))
```

and it is clear that our intuition holds. We can then compute discriminability, expecting a high value since the observations from the same random variable are so much closer than observations not from the same random variable:

```{r}
print(sprintf("Discriminability: %.2f", discr.discr(D, c(1,1,2,2))))
```

and we get perfect discriminability of 1.
