---
title: "Discriminability Statistic"
author: "Eric Bridgeford"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Discriminability}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Discriminability Statistic

In this notebook, we will discuss the basics of the discriminability statistic in a real dat simulated and real data context, and demonstrate some of the useful plots that users may want to visualize the results with.

## Simulated Data

Here, we assume that we have 5 independent sources of a measurement, and take 10 measurements from each source. Each measurement source $i = 1:n$ is a random variable where measurements taken at the source for the $j^{th}$ dimension is $X_{ij} \sim \mathcal{N}(i, 0.5)$.

```{r}
library(MGC)
require(latex2exp)

nsrc <- 5
nobs <- 10
d <- 20
set.seed(12345)
src_id <- array(1:nsrc)
labs <- sample(rep(src_id, nobs))
dat <- t(sapply(labs, function(lab) rnorm(d, mean=lab, sd=0.5)))
discr.discr(dat, labs)  # expect high discriminability since measurements taken at a source have the same mean and sd of only 0.5
```

we may find it useful to view the distance matrix, ordered by source label, to show that objects from the same source have a lower distance than objects from a different source:

```{R, fig.width=6, fig.height=4}
Dx <- as.matrix(dist(dat[sort(labs, index=TRUE)$ix,]), method='euclidian')
mgc.plot.plot_matrix(Dx, xlab="Sorted by Source", ylab="Sorted by Source", title="Distance Matrix for Simulated Data", legend.name="dist(x, y)")
```

as we can see, the ordering of the data elements does not matter, and users can pass in the data as either an $[n, d]$ array, or a $[n, n]$ distance matrix:

```{r}
discr.stat(Dx, sort(labs))
```
## Real Data

Below, we show how discriminability might be used on real data, by demonstrating its usage on the first $4$ dimensions of the `iris` dataset, to determine the relationship between the flower species and the distances between the different dimensions of the iris dataset (sepal width/length and petal width/length):

```{r, fig.width=6, fig.height=4}
Dx <- as.matrix(dist(iris[sort(as.vector(iris$Species), index=TRUE)$ix,c(1,2,3,4)]))

mgc.plot.plot_matrix(Dx, xlab="Sorted by Species", ylab="Sorted by Species", title="Distance Matrix for Iris Dataset", legend.name=TeX("$dist(x, y)$"))
```

we expect a high discriminability since the within-species relationship is clearly strong (the distances are low for same-species):

```{r}
discr.stat(iris[,c(1,2,3,4)], as.vector(iris$Species))
```

which is reflected in the high discriminability score.
