---
title: "Discriminability Statistic"
author: "Eric Bridgeford"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{discriminability}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Discriminability Statistic

In this notebook, we will discuss the basics of the discriminability statistic in a real dat simulated and real data context, and demonstrate some of the useful plots that users may want to visualize the results with.

```{r}
library(mgc)
library(reshape2)
library(ggplot2)

plot_mtx <- function(Dx, main.title="Distance Matrix", xlab.title="Sample Sorted by Source", ylab.title="Sample Sorted by Source") {
  data <- melt(Dx)
  ggplot(data, aes(x=Var1, y=Var2, fill=value)) +
    geom_tile() +
    scale_fill_gradientn(name="dist(x, y)",
                         colours=c("#f2f0f7", "#cbc9e2", "#9e9ac8", "#6a51a3"),
                         limits=c(min(Dx), max(Dx))) +
    xlab(xlab.title) +
    ylab(ylab.title) +
    theme_bw() +
    ggtitle(main.title)
}
```

## Simulated Data

Here, we assume that we have 5 independent sources of a measurement, and take 10 measurements from each source. Each measurement source `i` has measurements sampled from `N(i, I_d)` where `d=20`.

```{r}
nsrc <- 5
nobs <- 10
d <- 20
set.seed(12345)
src_id <- array(1:nsrc)
labs <- sample(rep(src_id, nobs))
dat <- t(sapply(labs, function(lab) rnorm(d, mean=lab, sd=1)))
discr.stat(dat, labs)  # expect high discriminability since measurements taken at a source have the same mean and sd of only 1
```

we may find it useful to view the distance matrix, ordered by source label, to show that objects from the same source have a lower distance than objects from a different source:

```{R, fig.width=5, fig.height=4}
Dx <- as.matrix(dist(dat[sort(labs, index=TRUE)$ix,]), method='euclidian')
plot_mtx(Dx)
```

as we can see, the ordering of the data elements does not matter, and users can pass in the data as either an $[n, d]$ array, or a $[n, n]$ distance matrix:

```{r}
discr.stat(Dx, sort(labs))
```
## Real Data

Below, we show how discriminability might be used on real data, by demonstrating its usage on the first $4$ dimensions of the `iris` dataset, to determine the relationship between the flower species and the distances between the different dimensions of the iris dataset (sepal width/length and petal width/length):

```{r, fig.width=5, fig.height=4}
Dx <- as.matrix(dist(iris[sort(as.vector(iris$Species), index=TRUE)$ix,c(1,2,3,4)]))

plot_mtx(Dx)
```

we expect a high discriminability since the within-species relationship is clearly strong (the distances are low for same-species):

```{r}
discr.stat(iris[,c(1,2,3,4)], as.vector(iris$Species))
```

which is reflected in the high discriminability score.
