---
title: "MGC Simulations"
author: "Eric Bridgeford"
date: "November 2, 2017"
output: html_document
---

# MGC Simulations

In this experiment, we will take a look at the 20 canonical simulations analyzed in the [MGC Paper repository](https://github.com/neurodata-papers/MGC). For details on the simulations, please check out the appendix to the paper, which can be found on arXiv [MGC paper link](https://arxiv.org/abs/1609.05148), with available pseudo code.

# Data

# Simulations

In the below simulations, we let $X \sim Unif(-1, 1)$, and $x$ is $n$ samples from $X$.

```{r}
gen.coefs <- function(d) {
  A = as.array(1/1:d, dim=c(d, 1))
  return(A)
}

gen.x <- function(n, d) {
  return(t(replicate(n, runif(d, -1, 1))))
}
```

## Linear Relationship

### 1 Dimensional

```{r}

gen.y.linear <- function(x, A, eps) {
  er <- rnorm(n, mean=0, sd=1)
  return(x%*%A + eps*er)  # y = xA + nu  
}

n <- 100
d <- 1
xld <- gen.x(n, d)
Ald <- as.vector(1)
ylinld <- gen.y.linear(xld, A, 0.2)
```

### High Dimensional

```{r}

```

## Exponential

## Cubic

## Joint Normal

## Step Function

## Quadratic

## W shape

## Spiral

## Bernoulli

## Logarithmic

## Fourth Root

## Circle

## Ellipse

## Diamond

## Multiplicative

## Independence

# Comparison

## 1 Dimensional

## High Dimensional

# Geometries
